Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula

### 2024-02-07 output/paper/_si.qmd
Kept working on papers from MV v1.

### Description of chemicals

@@

### Description of metabolites

@@

### Description of chemicals

@@

### Description of metabolites

>>>>

### Description of endocrine disruptors

@@

### Description of glucocorticosteroids

::: {#supptbl-met-desc}
```{r met-desc}
desc_chems_mets$desc_mets
```
**Participants glucocorticosteroids concentrations, by cohort and overall (HELIX subcohort; 2013-2016).**
:::

@@

::: {#suppfig-flow-pop}
```{mermaid}
flowchart TB
  helixsc["HELIX subcohort\n(N = 1,301)"] --> edcs["HELIX data EDCs\n(N = 1,297)"]
  helixsc --> corts["HELIX data glucocorticosteroids\n(N = 1,004)"]
  edcs --> inter["HELIX data EDCs and glucocorticosteroids\n(N = 976)"]
  corts --> inter
  edcs -.-> rq1(["HRT-SE ~ EDCs"])
  inter -.-> rq2(["metabolites ~ EDCs"])
  inter -.-> rq3(["HRT-SE ~ metabolites"])
```
Flowchart describing thew sample size for each research question.
:::

### Description of endocrine disruptors

@@

### Description of glucocorticosteroids

### 2024-02-07 output/paper/manuscript.qmd
Kept working on papers from MV v1.

format:
  docx:
    # reference-doc: reference_document.docx
    page-width: 8.3
    citeproc: true
  # pdf:
  #   pdf-engine: lualatex
  #   number-sections: true
  #   toc: false
  #   keep-tex: false
  #   include-in-header:
  #     text: |
  #       \usepackage{lineno}
  #       \usepackage{rotating}
  #       \usepackage{float}
  #       \usepackage{typearea}
  #   include-before-body:
  #     text: |
  #       \linenumbers

@@

To assess the levels of glucocorticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute). The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020]. Measurements were available for 980 children of the \acr{helix} subcohort.

@@

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with DAGitty [@TextorvanderZanderGilthorpe:2016] and ggdag [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. Codebooks for the used covariates, by research question, are provided in @supptbl-codebook-1, @supptbl-codebook-2, and @supptbl-codebook-3.

@@

In total, 1,297 children were included in the study population. Of these, 976 had information also about the glucocorticosteroids. The sample consisted of 55% girls. The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children. At the time of visit, the median age of the children was 8.06 years. The children were mostly Caucasian (90%), and the largest minority were of Pakistani origin (6.2%). @tbl-pop-desc provides descriptive statistics for the outcome and covariates, by cohort and overall.

Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids, by cohort and overall, are presented in @tbl-edc-desc, @tbl-met-desc, and @tbl-met-new-desc.

>>>>

format:
  docx:
    # reference-doc: reference_document.docx
    page-width: 8.3
    citeproc: true
    mermaid:
      theme: neutral
  # pdf:
  #   pdf-engine: lualatex
  #   number-sections: true
  #   toc: false
  #   keep-tex: false
  #   include-in-header:
  #     text: |
  #       \usepackage{lineno}
  #       \usepackage{rotating}
  #       \usepackage{float}
  #       \usepackage{typearea}
  #   include-before-body:
  #     text: |
  #       \linenumbers

@@

To assess the levels of glucocorticosteroids and their metabolites, LC-MS/MS analysis was applied at the Applied Metabolomics Research Group, IMIM (Hospital del Mar Medical Research Institute). The laboratory protocols for the analysis are described elsewhere [@MarcosRenauCasals:2014; @Gomez-GomezPozo:2020].

@@

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a \acr{dag} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].

@@

Of the 1,301 children of the \acr{helix} subcohort, 1,297 had measurements of the non-persistent \acr{edcs}. Measurements of the glucocorticosteroids were available for 1,004 children, of which 980 were matched to the \acr{helix} subcohort. Measurements of both non-persistent \acr{edcs} and glucocorticosteroids were available for 976 children of the subcohort. A flowchart describing the sample size for each research question is presented in @suppfig-flow-pop. The sample consisted of 55% girls. The median \acr{hitrtse} was 300 ms (\acr{iqr}, 231-368), with lower median values for EDEN, MOBA, and INMA, corresponding to the cohorts with older children. At the time of visit, the median age of the children was 8.06 years. The children were mostly Caucasian (90%), and the largest minority were of Pakistani origin (6.2%). @tbl-pop-desc provides descriptive statistics for the outcome and covariates, by cohort and overall.

Levels of unprocessed non-persistent \acr{edcs} and glucocorticosteroids are presented in @tbl-edc-desc, @tbl-met-new-desc, and @supptbl-met-desc.
